---
title: "PROBLEM SET 3"
author: "Jacob Jameson"
output:
  pdf_document: default
---
    
### IDENTIFICATION

(1) Your information

\color{blue}
Jacob Jameson
\color{black}


(2) Group Members (please list below the classmates you worked with on this problem set):

\color{blue}
Bohan Li, Jenna Rogers
\color{black}


(3)	Compliance with Harvard Kennedy School Academic Code


\color{blue}

I certify that my work in this problem set complies with the Harvard Kennedy School Academic Code

\color{black}



### Conceptual Questions 


### Data Questions


```{r, message=F, echo=F}
library(tidyverse)
library(foreign)
library(lmtest)
library(gridExtra)
library(haven)
library(labelled)
library(ggsci)
library(sandwich)
library(stargazer)
library(kableExtra)
library(AER)
library(lfe)
library(purrr)
library(broom)
library(vtable)
library(knitr)
suppressWarnings(library(kableExtra))
```


1.	Create summary statistics for povrate60, mort_age59_related_postHS and two other variables of your choice.  


```{r, results='hide', message=F, warning=F}
data <- read_csv('headstart.csv')


sumtable(data, 
         vars = c('povrate60', 'mort_age59_related_postHS',
                  'census1960_pop', 'census1990_pctblack'),
         out='latex')
```
\begin{table}[h] \centering \renewcommand*{\arraystretch}{1.1}\caption{Summary Statistics}\resizebox{\textwidth}{!}{
\begin{tabular}{lrrrrrrr}
\hline
\hline
Variable & N & Mean & Std. Dev. & Min & Pctl. 25 & Pctl. 75 & Max \\ 
\hline
povrate60 & 2804 & 37 & 15 & 15 & 24 & 47 & 93 \\ 
mort\_age59\_related\_postHS & 2785 & 2.3 & 5.7 & 0 & 0 & 2.8 & 136 \\ 
census1960\_pop & 2804 & 38964 & 117460 & 224 & 9133 & 33417 & 2664438 \\ 
census1990\_pctblack & 2806 & 0.091 & 0.15 & 0 & 0.0015 & 0.11 & 0.86\\ 
\hline
\hline
\end{tabular}
}
\end{table}


2.	 Create two balance tables for the 1960 and 1990 census variables (similar to Table I in the QJE article) comparing the county characteristics for counties above and below the poverty rate cutoff of 59.1984%. Add a column showing the estimated difference between the two sets of counties, and the p-value that this difference is statistically significant.








