---
title: "PROBLEM SET 2"
author: "Jacob Jameson"
date: "Due on Tuesday March 7, 2023"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
    
### IDENTIFICATION

(1) Your information

\color{blue}
Jacob Jameson
\color{black}



(2) Group Members (please list below the classmates you worked with on this problem set):

\color{blue}
Bohan Li, Jenna Rogers
\color{black}


(3)	Compliance with Harvard Kennedy School Academic Code

\color{blue}
I certify that my work in this problem set complies with the Harvard Kennedy School Academic Code
\color{black}



### Conceptual Questions 


1.	Read the paper. In a couple of sentences, explain the primary research question that the authors are trying to answer. What is the key logic behind the authors' answer? In a separate sentence, explain why policymakers should care about that question. Use non-technical jargon so that someone without statistical training could understand.



2.	The authors used an instrumental variable approach because they believed a naïve observational regression specification (regressing earnings on education) would be insufficient. What are two possible confounders (omitted variables) that would bias the results from this regression? Explain the mechanism of the omitted variable and use the omitted variable bias formula to argue whether it would lead to an understatement or overstatement of the true effect.





3.	What is/are the instrument(s) used by the authors in this study, and what are those instruments instrumenting for?



4.	Generally, what characteristics must an instrument have for it to be considered valid? 


a.	Explain these characteristics in words, and specific terms of the instrument(s) used in the paper.


b.	Explain these characteristics using random variables and potential outcomes. 




5.	Do you believe that the instrument(s) in the paper is/are truly exogenous? Why or why not? If so, provide a brief argument for this assumption. If not, provide an alternate mechanism for how the exogeneity assumption may be violated. 






6.	To assess whether the instrument is relevant, we can examine whether the instrument (quarter of birth) predicts the instrumentalized variable (compulsory schooling). In the following parts a - c, provide interpretations with concrete units, in a way a policymaker can understand.

a.	Explain how Table 1 is constructed, and give some intuition for the authors' choices.
 

b.	Interpret the coefficient of the first quarter for the outcome variables "Total years of education" and "High school graduate" of the 1930-1939 cohort.


c.	Why do the authors estimate coefficients for the bottom part of Table 1 ("College graduate", "Completed master’s degree", "Completed doctoral degree")? How do the results support the validity of their instrument? What assumption of the IV model are they addressing here?



7.	Consider Table III and Table IV. Provide a general formula and a basic intuition for the Wald estimator. How does it compare to the OLS estimate? What is the advantage of using TSLS?


8.	How would you construct a reduced form table? What would be the purpose thereof? What figure in the paper fulfills this purpose?



9.	Subsequent papers have found the instrument to be weak for some specifications of the paper. 


a.	Provide an intuition for the reason why weak instruments are problematic.


b.	Read the following, explain the intuition, and explain the implications for weak instruments.




c.	Optional: if you know what the bootstrap does, why does bootstrapping not solve the weak identification issue?



10.	If you were to write the paper today, how would you detect weak instruments, and what statistic would you use for inference?

Hint: you may want to refer to Andrews, Stock and Sun (2019).

a.	How is the effective F-statistic constructed?


b.	How are Anderson-Rubin confidence sets constructed?


### The Local Average Treatment Effect

11.	Explain the monotonicity assumption in the context of this study. What is required regarding the relationship between variables for monotonicity to be met, and is it reasonable to assume that defiers do not exist? In your explanation, be sure to touch on what does it means to be a defier in this study.


12.	Interpret the IV estimates in Table IV with appropriate units in the context of the study’s research question, treating them as a local average treatment effect. In your interpretation, clarify the population for which this local average treatment effect is identified (i.e., who are the compliers?).





13.	With 3-5 sentences, discuss the how you believe these results inform policy outside of the specified context.  In your discussion, be sure to touch on the problems specific to generalizing from instrumental variable findings and the particular scope conditions of this paper’s findings.






\newpage


## Data Analysis Questions (22 points + 2 extra points)


14.	Figure I can be thought of as a "graphical first-stage", as it shows the mean of completed years of education by quarter of birth for each year of birth between 1930 and 1940. Replicate Figure I. You can highlight the mean of years of education of those born in the first quarter (for each year between 1930 and 1940). 

Hint: you may want to create year of birth and quarter of birth dummies. They will also be useful for the following questions.



```{r, message=F, echo=F}
library(tidyverse)
library(foreign)
library(lmtest)
library(gridExtra)
library(haven)
library(labelled)
library(ggsci)
library(sandwich)
library(stargazer)
library(kableExtra)
library(AER)
library(lfe)
library(purrr)
library(broom)
library(knitr)
suppressWarnings(library(kableExtra))
```

```{r, message=F, warning=F}
data <- read_dta("data/AK91_1930_39.dta") %>%
  filter(YOB <= 39)

data$QOB <- as.factor(data$QOB) 

data %>%
  group_by(YOB, QOB) %>%
  summarise(mean.EDUC = mean(EDUC), 
            quarter.year = YOB + (as.numeric(QOB)/4) - 0.25) %>%
  ungroup() %>%
  unique() %>%
ggplot(aes(x = quarter.year, y = mean.EDUC)) +
  geom_path() +
  geom_point(aes(color=QOB), size = 3) + 
  geom_text(aes(label = paste(QOB)),
          vjust = 2.5, size = 3) +
  scale_x_continuous(breaks = seq(30, 40, by = 2)) +
  scale_y_continuous(breaks = seq(12.2, 13.2, by = .2)) +
  scale_color_manual(values = c("red", "black", "black", "black")) +
  labs(x = "Year of Birth", y = "Years of Completed Education",
       title = "Relationship between birth year and years of education",
       caption = paste("Figure I\n",
       "Years of Education and Season of Birth 1980 Census \n",
       "Note: Quarter of birth is listed below each observation")) +
  theme_classic() + 
  theme(plot.caption = element_text(
    color = "black", 
    size = 8,
    lineheight = 1.2, 
    hjust = 0.5,
    margin = margin(t = 20)),
    legend.position = "none")
```

\newpage

15.	Table I shows the relationship between quarter of birth and educational outcomes. Replicate the first row of Table I, i.e., find the coefficients of the first, second, and third quarter-of-birth dummies on total years of education. Your estimates do not need to be exactly the same as in the paper, but roughly of the same magnitude.


```{r, warning=F, message=F}
data <- within(data, QOB <- relevel(QOB, ref = 4))

ols <- tidy(lm(formula = paste0('EDUC', " ~ QOB"), data = data))

table <- data.frame(
    `Outcome variable` = character(),
    `Birth cohort` = character(),
    Mean = double(),
    I = double(),
    II = double(),
    III = double(),
    `F-test` = character())

table <- rbind(table, data.frame(
    `Outcome variable` = 'Total years of Education',
    `Birth cohort` = '1930-1939',
    Mean = round(mean(data$EDUC, na.rm = TRUE), 3),
    I = paste(round(ols$estimate[2], 3),
              paste0('(',round(ols$std.error[2], 3)),')'),
    II = paste(round(ols$estimate[3], 3), 
                paste0('(',round(ols$std.error[2], 3)),')'),
    III = paste(round(ols$estimate[4], 3), 
                 paste0('(',round(ols$std.error[2], 3)), ')'),
    `F-test` = summary(lm(formula = paste0('EDUC', " ~ QOB"), 
                          data = data))$fstat[[1]]))

kable(
    table, 
    caption = "Quarter-of-birth effect", 
    align = "c", 
    col.names = c("Outcome Variable", "Birth Cohort", 
                  "Mean", "I", "II", "III", "F-test")) %>%
  kable_styling(full_width = FALSE)
```




16.	Create a reduced form table. In other words, you want to regress the log of weekly earnings on the quarter of birth dummies (our instrument).  You may want to include year fixed effect.



```{r, results='hide'}
data$YOB <- as.factor(data$YOB)

reg.reduced.form <- felm(QOB ~ LWKLYWGE|0|0|0,data)
reg.reduced.form.fe <- felm(QOB ~ LWKLYWGE|YOB|0|0,data)

stargazer(reg.reduced.form, reg.reduced.form.fe, type = "html", 
          column.labels = c("Reduced Form 1 (OLS)", "Reduced Form  2 (FE)"))
```

<table style="text-align:center"><tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="2"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="2" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td colspan="2">Quarter of Birth</td></tr>
<tr><td style="text-align:left"></td><td>Reduced Form 1 (OLS)</td><td>Reduced Form 2 (FE)</td></tr>
<tr><td style="text-align:left"></td><td>(1)</td><td>(2)</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Log of weekly earnings</td><td>0.014<sup>***</sup></td><td>0.014<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.003)</td><td>(0.003)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">Constant</td><td>2.425<sup>***</sup></td><td></td></tr>
<tr><td style="text-align:left"></td><td>(0.017)</td><td></td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>329,509</td><td>329,509</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.0001</td><td>0.0002</td></tr>
<tr><td style="text-align:left">Adjusted R<sup>2</sup></td><td>0.0001</td><td>0.0002</td></tr>
<tr><td style="text-align:left">Residual Std. Error</td><td>1.112 (df = 329507)</td><td>1.112 (df = 329498)</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="2" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>

\newpage



17.	Table III reports OLS and Wald estimates of returns of education. Replicate both estimates for men born 1930-1939 (Panel B). 




```{r}
data$QOB_dummy <- ifelse(data$QOB == 1, 1, 0)

group1 <- data %>% filter(QOB == 1)
group2 <- data %>% filter(QOB != 1)

wage_1st <- mean(group1$LWKLYWGE)
se1 <- sd(group1$LWKLYWGE) / sqrt(nrow(group1))

wage_other <-  mean(group2$LWKLYWGE)
se2 <- sd(group2$LWKLYWGE) / sqrt(nrow(group2))

wage_diff <- wage_1st - wage_other
wage_err <- sqrt(se1^2 + se2^2)

educ_1st <- mean(group1$EDUC)
se1 <- sd(group1$EDUC) / sqrt(nrow(group1))

educ_other <- mean(group2$EDUC)
se2 <- sd(group2$EDUC) / sqrt(nrow(group2))

educ_diff <- educ_1st - educ_other
educ_err <- sqrt(se1^2 + se2^2)

# wald return to education
data$EDUC_pred <- lm(EDUC ~ QOB_dummy, data = data) %>% predict()
wald_rslt <- lm(LWKLYWGE ~ EDUC_pred, data = data)

# ols return to education
ols_rslt <- lm(LWKLYWGE ~ EDUC, data = data)

wald_est = coef(wald_rslt)["EDUC_pred"]
wald_err = sqrt(diag(vcovHC(wald_rslt, type = "HC1"))["EDUC_pred"])
ols_est = coef(ols_rslt)["EDUC"]
ols_err = sqrt(diag(vcovHC(ols_rslt, type = "HC1"))["EDUC"])

# create table
table <- data.frame(
    Variable = c("ln (wkly. wage)", "", 
                 "Education", "",
                 "Wald est. of return to education", "", 
                 "OLS return to education", ""),
    `(1) Born in 1st quarter of year` = 
      c(round(wage_1st,5), "", 
        round(educ_1st,5), "", 
        "", "", 
        "", ""),
    `(2) Born in 2nd, 3rd, or 4th quarter of year` = 
      c(round(wage_other,5), "", 
        round(educ_other,5), "", 
        "", "", 
        "", ""),
    `(3) Difference (std.error) (1) - (2)` = 
      c(paste(round(wage_diff,5), 
        paste0("(", round(wage_err ,5), ")")), "",
        paste(round(educ_diff,5), 
        paste0("(", round(educ_err, 5), ")")), "",
        paste(round(wald_est[[1]], 5), 
        paste0("(", round(wald_err[[1]], 5), ")")), "",
        paste(round(ols_est[[1]], 5), 
        paste0("(", round(ols_err[[1]], 5), ")")), "")
  )

kable(
    table,
    caption =  "Panel B: Wald Estimates for 1980 Census - Men Born 1930-1939",
    align = "c",
    col.names = c("", "(1) Born in 1st quarter of year", 
                  "(2) Born in 2nd, 3rd, or 4th quarter of year", 
                  "(3) Difference (std.error) (1) - (2)")) %>%
  kable_styling(full_width = FALSE)
```




18.	Table V reports different specifications of the TSLS for men born 1930-1939. Run TSLS similar as in Column 2 and Column 6. First, instrument education with a set of quarter-of-birth x year-of-birth dummies, and add year fix effects as control. Why do we want to include year fix effects?

Second, similarly to Column 6, instrument education with the same interaction dummies, and add regional fix effects, year fix effects, race, married status, and an urban dummy.



```{r, results='hide'}
ivreg.2 <- ivreg(LWKLYWGE ~ EDUC + YOB | QOB*YOB, data = data)

ivreg.6 <- ivreg(LWKLYWGE ~ EDUC + YOB + NEWENG + 
                   MIDATL + ENOCENT + WNOCENT + 
                   SOATL + ESOCENT + WSOCENT + MT +
                   RACE + SMSA + MARRIED | QOB*YOB, data = data)


varnames <- c("EDUC", "RACE", "SMSA", "MARRIED")
labels <- c("Years of Education", "Race (1 == black)", 
            "SMSA (1 == city center)", "Married (1 == married)")

birth_dummy <- c("Yes", "Yes")
region_dummy <- c("No", "Yes")

stargazer(ivreg.2, ivreg.6, type = "html", 
          coef.names = labels, 
          covariate.labels = labels, 
          covariate.include = varnames, 
          column.labels = c("Model 2 (TSLS)", "Model 6 (TSLS)"), 
          add.lines = list(c("9 Year of Birth dummies", birth_dummy[1], birth_dummy[2]), 
                           c("8 Region of Residence dummies", 
                             region_dummy[1], region_dummy[2])),
                    omit = c("QOB", "YOB", "NEWENG", "MIDATL",
                             "ENOCENT", "WNOCENT", "SOATL", "ESOCENT", 
                             "WSOCENT", "MT"))
```
<table style="text-align:center"><tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="2"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="2" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td colspan="2">LWKLYWGE</td></tr>
<tr><td style="text-align:left"></td><td>Model 2 (TSLS)</td><td>Model 6 (TSLS)</td></tr>
<tr><td style="text-align:left"></td><td>(1)</td><td>(2)</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Years of Education</td><td>0.089<sup>***</sup></td><td>0.068<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.016)</td><td>(0.024)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">Race (1 == black)</td><td></td><td>-0.411</td></tr>
<tr><td style="text-align:left"></td><td></td><td>(0.514)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">SMSA (1 == city center)</td><td></td><td>-0.783<sup>*</sup></td></tr>
<tr><td style="text-align:left"></td><td></td><td>(0.415)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">Married (1 == married)</td><td></td><td>0.486</td></tr>
<tr><td style="text-align:left"></td><td></td><td>(0.522)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">Constant</td><td>4.793<sup>***</sup></td><td>4.872<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.201)</td><td>(0.821)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">9 Year of Birth dummies</td><td>Yes</td><td>Yes</td></tr>
<tr><td style="text-align:left">8 Region of Residence dummies</td><td>No</td><td>Yes</td></tr>
<tr><td style="text-align:left">Observations</td><td>329,509</td><td>329,509</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.110</td><td>-0.082</td></tr>
<tr><td style="text-align:left">Adjusted R<sup>2</sup></td><td>0.110</td><td>-0.082</td></tr>
<tr><td style="text-align:left">Residual Std. Error</td><td>0.640 (df = 329498)</td><td>0.706 (df = 329487)</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="2" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>





19.	Now repeat the same exercise as the first 2LS2 in Question 18 (without additional controls, only year fix effects), but without using a built-in IV function – use only lm instead. Are your results the same as in the previous question?



```{r, results='hide'}
model_edu <- lm(EDUC ~ QOB*YOB, data = data)
data$educ_pred <- predict(model_edu)

model_lwklywge <- lm(LWKLYWGE ~ educ_pred + YOB, data = data)

stargazer(model_lwklywge, type = "html", 
          column.labels = c("Model 2 (TSLS) Only Using `lm`"), 
                    omit = c("QOB", "YOB"))
```


<table style="text-align:center"><tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="1" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td>LWKLYWGE</td></tr>
<tr><td style="text-align:left"></td><td>Model 2 (TSLS) Only Using `lm`</td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">educ_pred</td><td>0.089<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.017)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">Constant</td><td>4.793<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.213)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>329,509</td></tr>
<tr><td style="text-align:left">R<sup>2</sup></td><td>0.0001</td></tr>
<tr><td style="text-align:left">Adjusted R<sup>2</sup></td><td>0.0001</td></tr>
<tr><td style="text-align:left">Residual Std. Error</td><td>0.679 (df = 329498)</td></tr>
<tr><td style="text-align:left">F Statistic</td><td>3.709<sup>***</sup> (df = 10; 329498)</td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>


The results of the `lm()` and `ivreg()` models differ slightly because the `lm()` model does not directly address the endogeneity of `EDUC` like the `ivreg()` model does with instrumental variables. Using predicted values from a separate regression might not fully account for the endogeneity and could bias the estimates. For instance I can see that my coefficient is about the same, but the standard errors are different.




20.	For the purpose of this question, define treatment as completing high school (12 or more years of education), and the instrument as binary, with Zi=1 if one is born in the fourth quarter, and 0 otherwise.


```{r}
data$treatment <- ifelse(data$EDUC >= 12, 1, 0)
data$QOB_dummy.4 <- ifelse(data$QOB == 4, 1, 0)
```

a.	What is the share of the complier population?


```{r}
iv_reg <- ivreg(LWKLYWGE ~ treatment | QOB_dummy.4, data = data)

# Extract the LATE
late <- summary(iv_reg)$coefficients[2]

# Calculate the share of compliers
print(late / mean(data$treatment))
```


b.	What is the average untreated outcome among the never-takers?


```{r}
mean(ifelse(data$treatment == 0 & data$QOB_dummy.4 == 0, 
                  data$LWKLYWGE, NA), na.rm = TRUE)
```

c.	What is the average treated outcome among the always-takers?


```{r}
mean(ifelse(data$treatment == 1 & data$QOB_dummy.4 == 1, 
                  data$LWKLYWGE, NA), na.rm = TRUE)
```


d.	Is there selection into treatment? State the assumptions necessary for your conclusion.



### IVs in Your Own Research


21.	Think about a social relationship that would be best studied using an IV approach. Briefly state the research question and the main variables of interest in non-technical terms. 


The research question could be "What is the effect of nurse staffing levels on patient outcomes in hospitals?" The main variables of interest would be the nurse staffing levels (nurse-to-patient ratios) and the patient outcomes (re-admission rates, length of hospital stay).

Because nurse staffing levels may be endogenous (i.e., they may be correlated with unobserved factors that affect patient outcomes), an instrumental variable approach could be used to estimate the causal effect of nurse staffing levels on patient outcomes. A plausible instrument for nurse staffing levels could be a state-mandated minimum nurse-to-patient ratio. This instrument could be used to estimate the causal effect of nurse staffing levels on patient outcomes by focusing on the population of patients who are most likely to be affected by nurse staffing levels (those who are admitted to hospitals with staffing levels near the state-mandated minimum).



22.	Write out the empirical specification you would use and explain the equation. 


\[\text{Outcome}_i = \alpha + \beta\cdot\text{StaffingLevel}_i + \epsilon_i \]

\[\text{StaffingLevel}_i = \gamma\cdot\text{MinimumRatio}_i + \delta_i\]

In this specification, $\text{Outcome}_i$ represents the patient outcome for patient $i$, $\text{StaffingLevel}_i$ represents the nurse staffing level for patient $i$, and $\text{MinimumRatio}_i$ represents the state-mandated minimum nurse-to-patient ratio for the hospital where patient $i$ is admitted. The coefficient $\beta$ represents the estimated effect of a one-unit increase in nurse staffing levels on patient outcomes, holding all other factors constant. The coefficient $\gamma$ represents the estimated effect of a one-unit increase in the state-mandated minimum ratio on nurse staffing levels, holding all other factors constant. The error terms $\epsilon_i$ and $\delta_i$ capture the unobserved factors that affect patient outcomes and nurse staffing levels, respectively.

To estimate the causal effect of nurse staffing levels on patient outcomes, an instrumental variable approach can be used where the first-stage regression of $\text{StaffingLevel}_i$ on $\text{MinimumRatio}_i$ is used to obtain predicted values of nurse staffing levels. The predicted values are then used as the instrumental variable in the second-stage regression of $\text{Outcome}_i$ on the predicted values of $\text{StaffingLevel}_i$.





23.	If you clustered your standard errors or included fixed effects, explain why these methods reduced the likelihood of bias in your results (and if applicable, in which direction). If you did not, explain why these methods were not appropriate in your setting. 











